---
title: "Bootstrap confidence intervals"
author: "Becky Tang"
date: "2024-04-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Non-parametric

### Bootstrap CI for median

```{r}
data("faithful")
x <- faithful$eruptions * 60
n <- length(x)
samp_med <- median(x)
B <- 5000
delta_star <- rep(NA, B)
for(i in 1:B){
  xstar <- sample(x, size = n, replace = T)
  delta_star[i] <- median(xstar) - samp_med
}

# bootstrap distribution of delta
hist(delta_star)

# (approximate) bootstrap CI
ci <- samp_med - quantile(delta_star, c(0.975, 0.025))
names(ci) <- c("2.5", "97.5")
ci
```

`r ci`

The approximate 95% bootstrap CI for the median eruption time of Old Faithul is [`r ci[1]`, `r ci[2]`].

### Bootstrap CI for proportion

```{r}
x <- c(rep("AA", 342), rep("Aa", 500), rep("aa", 187))
n <- length(x)
B <- 1000
theta_hat <- (2*sum(x == "AA" ) + sum(x == "Aa"))/(2*n)

delta_star <- rep(NA, B)
for(i in 1:B){
  x_star <- sample(x, size = n, replace = T)
  delta_star[i] <- (2*sum(x_star == "AA" ) + sum(x_star == "Aa"))/(2*n) - theta_hat
}

theta_hat - quantile(delta_star, c(0.975, 0.025))
```

## Parametric

Returning to the Old Faithful example: suppose we assume that the eruption times are Normal (even though the data clearly show that they are not...). For a parametric bootstrap, we will take repeated samples from a Normal distribution with mean and variance estimated from the observed data, then proceed as we did in the nonparametric bootstrap:

```{r}
x <- faithful$eruptions * 60
n <- length(x)
samp_mean <- mean(x)
samp_sd_mle <- sd(x)*sqrt((n-1)/n)
samp_med <- median(x)
B <- 5000
delta_star <- rep(NA, B)
for(i in 1:B){
  xstar <- rnorm(n, samp_mean, samp_sd_mle)
  delta_star[i] <- median(xstar) - samp_med
}
hist(delta_star)

# (approximate) bootstrap CI
ci <- samp_med - quantile(delta_star, c(0.975, 0.025))
names(ci) <- c("2.5", "97.5")
ci
```

