---
title: "Empirical Bootstrap"
author: "Becky Tang"
date: "2024-04-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 5, fig.height = 3)
```

## Example 1: bootstrapping for a mean

Consider the following sample of eruption times (in seconds) of the Old Faithful geyser:

```{r}
# obtain data
data("faithful")
x <- faithful$eruptions * 60
hist(x)
```

Perhaps we'd like to obtain a bootstrap distribution for the mean eruption time.

```{r}
n <- length(x)
B <- 5000
delta_star <- rep(NA, B)
for(i in 1:B){
  # if size isn't specified in sample(), defaults to length of x!
  xstar <- sample(x, replace = T)
  delta_star[i] <- mean(xstar) 
}

# bootstrap distribution of means
hist(delta_star)

# bootstrap estimate of bias
mean(delta_star) - mean(x)
```

## Example 2: bootstrapping for MSE

The coefficient of variation of a distribution is the quantity $\frac{\sigma}{\mu}$, where $\sigma$ and $\mu$ are the standard deviation and mean of the distribution, respectively. Let us generate some Poisson data and obtain estimates of the mean squared error (MSE) of the estimator $\frac{s}{\bar{x}}$ where $s$ is the sample standard deviation and $\bar{x}$ is the sample mean.

```{r}
# generate some Poisson data
lambda_true <- 1
n <- 20
x <- rpois(n, lambda_true)
```

An estimate for coefficient variation from the data is:

```{r}
cv_hat <- sd(x)/mean(x)
cv_hat
```

### Bootstrap estimate

```{r}
B <- 5000
delta_star <- rep(NA, B)
for(i in 1:B){
  xstar <- sample(x, replace = T)
  delta_star[i] <- sd(xstar)/mean(xstar)
}
hist(delta_star)

# bootstrap estimate of MSE
mean((delta_star - cv_hat)^2)
```

### "Monte Carlo" estimate

Since we know the true distribution, we don't need to bootstrap to estimate the MSE of $\frac{s}{\bar{x}}$ for $\frac{\sigma}{\mu}$. That is, rather than resampling from the original observation `x`, we can actually simulate new data sets from the Poisson distribution. Then we can compare the estimates to the true coefficient of variation, which for a $\text{Poisson}(\lambda)$ distribution is $\frac{\sqrt{\lambda}}{\lambda}$. 

```{r}
delta_hat <- rep(NA, B)
for(i in 1:B){
  x_new <- rpois(n, lambda_true)
  delta_hat[i] <- sd(x_new)/mean(x_new)
}
hist(delta_hat)

cv_true <- sqrt(lambda_true)/lambda_true
mean((delta_hat - cv_true)^2)
```
