---
title: "STAT 311: Problem Set 4 (R)"
author: "Your name here"
date: "03/13/2024"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Note: if you have never installed tidyverse, you will have to do so now. In the CONSOLE at the bottom, paste in and run the the following code (without the #):
# install.packages("tidyverse")

# You will see some text running for a bit, but eventually everything should be successful! Then run this chunk again.
library(tidyverse)
```

## Question 1

We will now simulate some German Tank data, obtain estimates of the number of tanks, and compare the estimators.

a)  Create three variables:
    -   `n_sim` for the number of simulations (set this to be at least 1000)

    -   `N_true` for the true number of tanks (make this 100); and

    -   `n` for the number of tanks captured (set this to 5)

```{r}
# code for part (a)

```

b)  Write a function called `sim_tanks` that takes in two arguments: the true number of tanks and the number of tanks captured. The function should simulate the German tank scenario by randomly sampling the tanks without replacement. The function should return a **vector** of length three of the estimates under $\hat{N}_{1}$, $\hat{N}_{2}$, and $\hat{N}_{3}$, as defined on the homework.

```{r}
# code for part (b)

```

c)  Using the `replicate()` function, simulate this problem for `n_sim` iterations. Then use the `t()` function to take the *transpose* of the output such that you have 3 columns and `n_sim` rows. Be sure to store your output into a variable!

*In the Console, take a look at the output. Make sure you understand what each row/column represents!*

```{r}
set.seed(311)
# code for part (c)

```

d)  Using your simulations from (c), obtain the empirical mean, variance, and MSE of $\hat{N}_{1}$, $\hat{N}_{2}$, and $\hat{N}_{3}$. You can access a specific column $j$ of matrix $M$ using the following code: `M[,j]`. Similarly, to access a specific row: `M[j,]`. Comment on how the values compare across the estimates. Based on what you've found, which estimator(s) appear to be unbiased? Which estimator(s) appear to be "best"?

```{r}
# means


# variances


# MSEs

```

**Solution**:

e)  Lastly, we will visualize the simulations of the estimates. Run the first line of code below and then take a look at what `df` looks like by typing `View(df)` in the Console.

Then fill in the remaining code and then set `eval = TRUE` before knitting. Briefly comment on what you notice.

```{r eval = FALSE}
df <- data.frame(R) %>%
  rename("N1" = 1, "N2" = 2, "N3" = 3) %>%
  pivot_longer(cols = 1:3, names_to = "estimator", values_to = "estimate") 

df %>%
  ggplot(., aes(x = ______, # name of variable in df that represents the values of the estimates
                fill = _____ ))+ # name of variable in df that represents which of the N_i
  geom_histogram(alpha = 0.5, binwidth = 2) +
  geom_vline(xintercept = ____ ) # variable representing the true total number of tanks
```

**Solution**:
